Page not found

### 2023-06-21 manuscript/chapters/4_locality.qmd
plop
goodbye i will now sleep for 3 days

Therefore, if $h_U(\vec v, t)$ solves the neural field equation in $[0,1]^2$, then $\tilde h(t) = h_U(t) \circ S^{-1}$ solves the equivalent neural field equation in $[0,1]$. The same reasoning applies in reverse: if $\tilde h(\alpha, t)$ solves the neural field equation in $[0,1]$, then $h_U(t) = \tilde h \circ S$ solves the neural field in $[0,1]^2$.

@@

```{python}
#| label: fig-distance-2d-1d
#| fig-cap: Distance between the first population and all other populations, in 1D and 2D. The dotted line has a slope $\sqrt 2$.
#| echo: false

@@

With this view, we understand that the coarse-grained performed in @sec-coarse-graining was making use of that fact that locality helps by bounding the variation inside of each averaged bin. Therefore, we would like for mathematically formulate that on average, the variation inside of each bin before coarse-graining is small. @fig-sketch-binned-variation gives an intuition of this. Even though the $[0,1]$ embedding might be discontinuous in places, on average nearby neurons should exhibit the same activity levels.

![Intuition behind the average binned variation. To simplify the visualization, activity levels are used in lieu of the 2D positions. 4 bins are shown with a different colour. The horizontal line represents the average activity in the bin. From top to bottom: random, column, Z mapping.](figures/fig-variation.svg){width=600px #fig-sketch-binned-variation}

@@

The definition of locality can essentially be seen as a weakened version of continuity, or "continuity on average". A continuous function $f$ evidently verifies $V_n(f) \xrightarrow{n \to \infty} 0$, since small neighbourhoods are mapped to small neighbourhoods. The reverse is obviously not true, since for instance the metric $V_n$ vanishes when applied to a step function, which is only piecewise continuous.

@@

Before we move to analytical derivations applying the average variation to mappings, we can first check numerically if it gives the expected results; that is $V_n$ vanishes a $n\to\infty$ for local mappings, whereas $V_n$ is lower bounded by a constant for non-local mappings. @fig-vnum shows this is indeed the case, rejoice!

```{python}
#| label: fig-vnum
#| fig-cap: (Numerator of the) average binned variation as a function of $n$. The local mappings have a growth slower than the upper bound $2\times 2^n$.
#| echo: false

@@

$$
\begin{aligned}
Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3} \cdots i_1}_\text{$n/2$ bits} b_n b_{n+2} \cdots b_{2m-1} \\
Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4} \cdots i_0}_\text{$n/2$ bits} b_{n+1} b_{n+3} \cdots b_{2m}
\end{aligned}
$$

If $n$ is odd, we have :

$$
\begin{aligned}
Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3} \cdots i_0}_\text{$(n+1)/2$ bits} b_n b_{n+2} \cdots b_{2m-1} \\
Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4} \cdots i_1}_\text{$(n-1)/2$ bits} b_{n+1} b_{n+3} \cdots b_{2m}
\end{aligned}
$$

We only consider even $n$ in the following, as the proof remains very similar.

@@

Putting both together, we can write :

$$
\begin{aligned}
V_n(C^{-1}) &= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^{-1}_1(\alpha^\prime) - C_1^{-1}(\alpha)| + |C^{-1}_2(\alpha^\prime) - C_2^{-1}(\alpha)| \\
&\geq
\begin{aligned}[t]
	\frac{1}{2^n}\sum_{i=1}^{2^n}&
	\sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^{-1}_1(\alpha^\prime) - C_1^{-1}(\alpha)|\\
	&+ \sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^{-1}_2(\alpha^\prime) - C_2^{-1}(\alpha)|
\end{aligned}\\
&\geq \frac 12 + \bO(2^{-n})
\end{aligned}
$$

@@

$$
V_n(C^{-1}) > \frac 12 \implies \lim_{n \to \infty} V_n > \frac 12
$$

@@

In @sec-Vn, we have introduced the average variation metric $V_n$. Assuming that $V_n \xrightarrow{n\to\infty} 0$ (which expresses the locality property), does this imply that the numerical estimation of the integral in the neural field equation in $[0,1]$ converges to its true value?

The intuition behind the proof of convergence is that the variation, coming from the coarse-graining, inside each numerical bin vanishes as we increase the number of iterations $n$, even though the the neural field $\tilde{h}$ and the connectivity kernel $\tilde{w}$ become fractal.

@@

$$
\begin{aligned}
|\mathrm{NI}_n - \mathrm{AI}| &\leq \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \Delta(\beta) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
\text{where}\ \Delta(\beta) &= \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) - \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t))\\
&= w_U(\alpha, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\alpha, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t))
\end{aligned}
$$

The main source of variance inside the integrand $\Delta(\beta)$ is the term $S^{-1}$, since $w_U$ and $h_U$ are "nice" functions. We express this regularity by the added assumption that $w_U(\alpha, \vec u) \phi(h_U(t, \vec u))$ is Lipschitz in $\vec u$, with Lipschitz constant $L(\alpha)$, and $\norm{\alpha}$ the corresponding norm on $[0,1]^2$. We argue this is not a strong assumption, because this is just expressing the regularity of the integrand in the integral over $[0,1]^2$.

A more pragmatic justification of this is to notice that the recurrent currents are expressed as:

$$
I^\text{rec}_U(\vec v) = w_U(\vec v, \vec u) \phi(h_U(t, \vec u)),
$$

and that in @sec-cycling-nf, numerical simulations showed that $I^\text{rec}_U(\vec v)$ is a well-behaved function.

@@

$$
\begin{aligned}
\Delta(\beta) &= w_U(\alpha, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\alpha, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)) \\
&\leq L(\alpha) \norm{S^{-1}(\beta_i) - S^{-1}(\beta)}
\end{aligned}
$$

@@

$$
\begin{aligned}
|\mathrm{NI}_n - \mathrm{AI}| &\leq L(\alpha) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
&\leq L(\alpha) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \left[\lambda \circ S^{-1}\right](\d \beta^\prime) \right| \\
&= L(\alpha) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \\
&\leq L(\alpha) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta, \beta^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} \norm{S^{-1}(\beta^\prime) - S^{-1}(\beta)} \\
&= L(\alpha) V_n \\
&\xrightarrow{n \to \infty} 0
\end{aligned}
$$

@@

$$
\begin{aligned}
Z_p^n(\vec{v^{(n)}}) &= \sum_{k=1}^{n} \sum_{\mu=1}^p b^{\mu}_k 2^{-(p(k-1)+\mu)} \\
&= 0.b^1_1 b^2_1 \cdots b^p_1 b^1_2 b^2_2 \cdots b^p_{n-1} b^1_n b^2_n \cdots b^p_n
\end{aligned}
$$

The same arguments apply to show that this generalized Z-mapping converges pointwise to $Z_p$, has the locality property, and that $V_n(Z_p) \xrightarrow{n\to\infty} 0$. Then the proof of convergence, albeit more laborious, can be repeated using the same procedure.

>>>>

Therefore, if $h_U(\vec v, t)$ solves the neural field equation in $[0,1]^2$, then $\tilde h(\alpha, t) = (h_U(\cdot, t) \circ S^{-1})(\alpha)$ solves the equivalent neural field equation in $[0,1]$. The same reasoning applies in reverse: if $\tilde h(\alpha, t)$ solves the neural field equation in $[0,1]$, then $h_U(\vec v, t) = (\tilde h(\cdot, t) \circ S)(\vec v)$ solves the neural field in $[0,1]^2$.

@@

```{python}
#| label: fig-distance-2d-1d
#| fig-cap: Distance between the first population and all other populations, in 1D and 2D. The dotted line has a slope $\sqrt 2$. This figure is generated using $n=6$.
#| echo: false

@@

With this view, we understand that the coarse-graining performed in @sec-coarse-graining makes use of that fact that locality helps by bounding the variation inside of each averaged bin. Therefore, we would like for mathematically formulate that on average, the variation inside of each bin before coarse-graining is small. @fig-sketch-binned-variation gives an intuition of this. Even though the $[0,1]$ embedding might be discontinuous in places, on average nearby neurons should exhibit the same activity levels.

![Intuition behind the average binned variation. To simplify the visualization, activity levels are used in lieu of the 2D positions. 4 bins are each shown with a different colour. The horizontal line represents the average activity in each bin. From top to bottom: Random, Column, Z-mapping.](figures/fig-variation.svg){width=600px #fig-sketch-binned-variation}

@@

The definition of locality can essentially be seen as a weakened version of continuity, or "continuity on average". A continuous function $f$ evidently verifies $V_n(f) \xrightarrow{n \to \infty} 0$, since small neighbourhoods are mapped to small neighbourhoods. The reverse is not true, since for instance the metric $V_n$ vanishes when applied to a step function, which is only piecewise continuous.

@@

Before we move to analytical derivations applying the average variation to mappings, we can first check numerically if it gives the expected results; that is $V_n$ vanishes when $n\to\infty$ for local mappings, whereas $V_n$ should be lower bounded by a constant for non-local mappings. @fig-vnum shows this is indeed the case, rejoice!

```{python}
#| label: fig-vnum
#| fig-cap: (Numerator of the) average binned variation as a function of $n$. The Z-mapping has a growth slower than the upper bound $2\times 2^n$. The Column mapping grows at the same rate as the upper bound, and the Random mapping saturates it.
#| echo: false

@@

$$
\begin{aligned}
Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3} \cdots i_1}_\text{$n/2$ bits} b_n b_{n+2} \cdots b_{2m-1} \\
Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4} \cdots i_0}_\text{$n/2$ bits} b_{n+1} b_{n+3} \cdots b_{2m}.
\end{aligned}
$$

If $n$ is odd, we have:

$$
\begin{aligned}
Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3} \cdots i_0}_\text{$(n+1)/2$ bits} b_n b_{n+2} \cdots b_{2m-1} \\
Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4} \cdots i_1}_\text{$(n-1)/2$ bits} b_{n+1} b_{n+3} \cdots b_{2m}.
\end{aligned}
$$

We only consider even $n$ in the following, as the proof for the latter remains very similar.

@@

Putting both together, we can write:

$$
\begin{aligned}
V_n(C^{-1}) &= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^{-1}_1(\alpha^\prime) - C_1^{-1}(\alpha)| + |C^{-1}_2(\alpha^\prime) - C_2^{-1}(\alpha)| \\
&\geq
\begin{aligned}[t]
	\frac{1}{2^n}\sum_{i=1}^{2^n}&
	\sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^{-1}_1(\alpha^\prime) - C_1^{-1}(\alpha)|\\
	&+ \sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^{-1}_2(\alpha^\prime) - C_2^{-1}(\alpha)|
\end{aligned}\\
&\geq \frac 12 + \bO(2^{-n}).
\end{aligned}
$$

@@

$$
V_n(C^{-1}) > \frac 12 \implies \lim_{n \to \infty} V_n(C^{-1}) > \frac 12
$$

@@

In @sec-Vn, we have introduced the average variation metric $V_n$. Assuming that $V_n(S^{-1}) \xrightarrow{n\to\infty} 0$ (which expresses the locality property), does this imply that the numerical estimation of the integral in the neural field equation in $[0,1]$ converges to its true value?

The intuition behind the proof of convergence is that the variation, coming from the coarse-graining, inside each numerical bin vanishes as we increase the number of iterations $n$, even though the neural field $\tilde{h}$ and the connectivity kernel $\tilde{w}$ become fractal.

@@

$$
\begin{aligned}
|\mathrm{NI}_n - \mathrm{AI}| &\leq \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \Delta(\beta) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
\text{where}\ \Delta(\beta) &= \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) - \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t))\\
&= w_U(\alpha, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\alpha, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)).
\end{aligned}
$$

The main source of variance inside the integrand $\Delta(\beta)$ is the term $S^{-1}$, since $w_U$ and $h_U$ are "nice" functions. We express this regularity by the added assumption that $w_U(\alpha, \vec u) \phi(h_U(t, \vec u))$ is Lipschitz in $\vec u$, with Lipschitz constant $L(\alpha)$, and $\norm{\alpha}$ is the corresponding norm on $[0,1]^2$. We argue this is not a strong assumption, because this is just expressing the regularity of the integrand in the integral over $[0,1]^2$.

A more pragmatic justification of this is to notice that the "density of recurrent currents" is expressed as:

$$
I^\text{rec}_U(\vec v, \vec u) = w_U(\vec v, \vec u) \phi(h_U(t, \vec u)),
$$

and that in @sec-cycling-nf, numerical simulations showed that $I^\text{rec}_U(\vec v, \vec u)$ (the integrand of the integral on $[0,1]^2$) is a well-behaved function.

@@

$$
\begin{aligned}
\Delta(\beta) &= w_U(\alpha, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\alpha, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)) \\
&\leq L(\alpha) \norm{S^{-1}(\beta_i) - S^{-1}(\beta)}.
\end{aligned}
$$

@@

$$
\begin{aligned}
|\mathrm{NI}_n - \mathrm{AI}| &\leq L(\alpha) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
&\leq L(\alpha) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \left[\lambda \circ S^{-1}\right](\d \beta^\prime) \right| \\
&= L(\alpha) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \\
&\leq L(\alpha) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta, \beta^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} \norm{S^{-1}(\beta^\prime) - S^{-1}(\beta)} \\
&= L(\alpha) V_n(S^{-1}) \\
&\xrightarrow{n \to \infty} 0
\end{aligned}
$$

@@

$$
\begin{aligned}
Z_p^n(\vec{v^{(n)}}) &= \sum_{k=1}^{n} \sum_{\mu=1}^p b^{\mu}_k 2^{-(p(k-1)+\mu)} \\
&= 0.b^1_1 b^2_1 \cdots b^p_1 b^1_2 b^2_2 \cdots b^p_{n-1} b^1_n b^2_n \cdots b^p_n.
\end{aligned}
$$

The same arguments apply to show that this generalized Z-mapping converges pointwise to $Z_p$, has the locality property, and that $V_n(Z_p^{-1}) \xrightarrow{n\to\infty} 0$. Then the proof of convergence, albeit more laborious, can be repeated using the same procedure.

### 2023-06-21 manuscript/chapters/5_conclusion.qmd
goodbye i will now sleep for 3 days

The use of the notion of regularity for the analysis of large-scale neuronal recordings is in its infancy @Stringer2019. This purely theoretical thesis gives a concrete example of an important conceptual fact: if we want to analyse large-scale neuronal recordings via nonlinear dimensionality reduction techniques (e.g. if we want to model neuronal activity as neural fields evolving on some abstract latent embedding space) the notion of dimensionality is meaningless without an associated notion of regularity. Indeed, our example shows that some neural field dynamics (neuronal population dynamics) can be equivalently expressed as dynamics on the embedding space $\R^2$ and the embedding space $[0,1]$ (but the regularity of the corresponding neural fields are different).

>>>>

The use of the notion of regularity for the analysis of large-scale neuronal recordings is in its infancy @Stringer2019. This purely theoretical thesis gives a concrete example of an important conceptual fact: if we want to analyse large-scale neuronal recordings via nonlinear dimensionality reduction techniques (e.g. if we want to model neuronal activity as neural fields evolving on some abstract latent embedding space), the notion of dimensionality is meaningless without an associated notion of regularity. Indeed, our example shows that some neural field dynamics (neuronal population dynamics) can be equivalently expressed as dynamics on the embedding space $\R^2$ and the embedding space $[0,1]$ (but the regularity of the corresponding neural fields are different).

### 2023-06-21 manuscript/index.qmd
goodbye i will now sleep for 3 days

Early recordings of cat's somatosensory and visual cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
The columnar organization of the cortex motivated the design of spatially structured models of neural population dynamics, namely neural field models, where the spatial dimensions corresponded to the two dimensions of the cortical sheet.
Later it has been shown that not only the physical position of the neurons can be used to parametrize their function, but also their location in more abstract "embedding spaces", for example a space of "neuron celltypes".

>>>>

Early recordings of cats' somatosensory and visual cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
The columnar organization of the cortex motivated the design of spatially structured models of neural population dynamics, namely neural field models, where the spatial dimensions corresponded to the two dimensions of the cortical sheet.
Later it has been shown that not only the physical position of the neurons can be used to parametrize their function, but also their location in more abstract "embedding spaces", for example a space of "neuron celltypes".
